---
title: "homework6"
author: "Michael Mila"
date: "2/28/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

## Read in data
```{r}
HAPPY <- readRDS("data/HAPPY.rds")
```

# Cleaning data

## Replacing IAP, DK, NA with "NA"
```{r}
HAPPY = replace(HAPPY, HAPPY == "IAP", NA)
HAPPY = replace(HAPPY, HAPPY == "DK", NA)
HAPPY = replace(HAPPY, HAPPY == "NA", NA)
HAPPY
```

## Changing "89 AND OLDER" to 89, and assuming the variable should be numeric
```{r}
HAPPY <- HAPPY %>% mutate(
  age = replace(AGE, AGE == "89 AND OLDER", 89),
  age = as.numeric(age)
) %>% select(-AGE)
str(HAPPY)
```

## Bring all levels of factor into a sensible order.

### Degree
```{r}
HAPPY <- HAPPY %>% mutate(
  degree = factor(tolower(DEGREE)),
  degree = factor(degree, levels=c("lt high school", "high school", "junior college", "bachelor", "graduate school"))
) %>% select(-DEGREE)

HAPPY %>% ggplot(aes(x = degree)) + geom_bar()
```


```{r}
HAPPY <- HAPPY %>% mutate(
  HAPPY = as.factor(HAPPY),
  SEX = as.factor(SEX),
  MARITAL = as.factor(MARITAL),
  FINRELA = as.factor(FINRELA),
  HEALTH = as.factor(HEALTH),
  PARTYID = as.factor(PARTYID)
)
str(HAPPY)
```

### Changing levels of marital status and finanical relationship
```{r}
HAPPY <- HAPPY %>% mutate(
  marital = factor(tolower(MARITAL)),
  marital = factor(marital, levels=c("never married", "seperated", "married", "divorced", "widowed"))
) %>% select(-MARITAL)

levels(HAPPY$marital)

HAPPY <- HAPPY %>% mutate(
  finrela = factor(tolower(FINRELA)),
  finrela = factor(finrela, levels=c("far below average", "below average", "average", "above average", "far above average"))
) %>% select(-FINRELA)

levels(HAPPY$finrela)
```

str(HAPPY)
levels(HAPPY$HEALTH)

### Changing of levels of health to make it from poor to excellent

```{r}
HAPPY <- HAPPY %>% mutate(
  health = factor(tolower(HEALTH)),
  health = factor(health, levels=c("poor", "fair", "good", "excellent"))
) %>% select(-HEALTH)

levels(HAPPY$health)
```

str(HAPPY)
levels(HAPPY$PARTYID)
